---
title: "PSYC121: Week 6 Lab"
output: 
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
---

```{r setup, include=FALSE, exercise.startover = TRUE}
library(tidyverse)
library(learnr)
load("data_clean.RData") # loads as "data"
knitr::opts_chunk$set(echo = FALSE)
```

## Selecting and viewing datasets

### The questionnaire data

We are going to take a look at the data from the questionnaire you completed in Week 1. The data is loaded into the exercises here, so we can look at it by simply calling it with the name "data":

```{r Qdata, exercise = TRUE}
data

```

This is quite a big dataset to work with. We can see the size of it in a couple of ways. First we can use the "dim" function to see the number of rows and the number of columns. We can also use the "colnames" function to see a list of the column names for the data:


```{r dimensions, exercise = TRUE}
dim(data) # give the dimensions (rows,colums)
colnames(data) # list the column names
```

## Selecting columns

Let's just take a few of these columns to work with in our exercises. To do this we can use the **"select"** command, which is part of dplyr/tidyverse. To use this, specify the dataset you want to use (data), and then the columns you want to select. To specify a list of columns, use the "c" function:

```{r select_1, exercise = TRUE}
select(data, c(id,age,countries_visited,stroop_control,stroop_compatible,stroop_incompatible))

```
There are a couple of things that make this command a bit easier to work with. First, you can put the column names on a new line, to create an easier to read list. Anything after a comma in a function can be put on a new line. Secondly, if you need several columns in a row, you can use the colon (:) between the first and last column you need:

```{r select_2, exercise = TRUE}
select(data, c(id,
               age,
               countries_visited,
               stroop_control:stroop_incompatible))

```

It's now your turn to try out the select command. Here you will select a new set of columns and create a new tibble from that selection. Include all of those listed in the last bit of code (you can copy that code if you like) but also add all of the data on social media. Store the result of this selection in a new variable called "data_short".

```{r select_3, exercise = TRUE}


```

```{r select_3-hint-1}

data_short <- select(data, )

```

```{r select_3-solution}

data_short <- select(data, c(id,
                             age,
                             countries_visited,
                             stroop_control:stroop_incompatible,
                             facebook_days:twitter_follow))

```

which version of data_short is produced?

```{r outputData, exercise = TRUE}

data_short

```


